---
title: "Lab3"
author: "Your Name"
date: "2024-09-19"
output: "github_document"
---

# Load Packages
```{r}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(psych)
library(bruceR)
library(ggsci)
library(see)

```

# Load Datasets
```{r}

lab3data <- read.csv("/Users/caseyzhang/Documents/GitHub/Lab3/lab3data.csv")

```


# Missing Data
```{r}

#Listwise deletion
list_lab3data<- drop_na(lab3data)

#pair wise deletion will only be used when you run the actual analysis using "na.rm = TRUE"

```

# Recode Variables: Changing Numeric Values to Characters
```{r}

lab3data %>% count(Gender)

#Solution 1
lab3data<- lab3data%>%
  mutate_at(c('Gender'),funs(str_replace(., "1", "Male")))

#Solution 2

#lab3data$Gender<-recode(lab3data$Gender, '1' = 'X', '2' = 'Y')


#Now recode 2 into Female
lab3data<- lab3data%>%
  mutate_at(c('Gender'),funs(str_replace(., "2", "Female")))



```

# Recode Variables: Changing Numeric Values to Different Numeric Values
```{r}

lab3data %>% count(RSE2)

lab3data$RSE2_R <- 5 - lab3data$RSE2

#check to see if the recoding was successful
lab3data %>% count(RSE2_R)

#Now do the same for the other items (5,6,8,9)
lab3data$RSE5_R <- 5 - lab3data$RSE5
lab3data$RSE6_R <- 5 - lab3data$RSE6
lab3data$RSE8_R <- 5 - lab3data$RSE8
lab3data$RSE9_R <- 5 - lab3data$RSE9

```


# Create Composites
```{r}

#Depending on the scale, sometimes you need to average across the items, other times you might need to sum 


#Fill in the items
lab3data <- lab3data %>%
  mutate(RSE = rowMeans(cbind(RSE1, RSE2_R, RSE3, RSE4, RSE5_R, RSE6_R, RSE7, RSE8_R, RSE9_R, RSE10)))

#Create another composite for SWL, but this time use rowSums
lab3data <- lab3data %>%
  mutate(SWL = rowSums(cbind(SWL1, SWL2, SWL3, SWL4, SWL5)))
```


# Get Summary Descriptive Statistics
```{r}

lab3data %>% 
  group_by(Gender) %>%
  dplyr::summarize(mean_RSE    = mean(RSE),
      mean_SWL    = mean(SWL),
      std_dev_RSE = sd(RSE),
      std_dev_SWL = sd(SWL),
      corr_RSE_SWL  = cor(RSE, SWL)
    )

```

# Q1: Descrptively, which group has a higher self-esteem? What about life satisfaction? Is the relationship between self-esteem and life satisfaction stronger for males or for females? 

Self-esteem is higher for males (3.12) than females (2.98). Life satisfaction is higher for females (16.29) than males (16.25). The relationship between self-esteem and life satisfaction is stronger for males (0.50) than females (0.36).


# Visualization of Data Distribution
```{r}

lab3data<- lab3data %>%
  drop_na(Gender)


ggplot(lab3data, aes(x = Gender, y = RSE)) + geom_violin(aes(fill=Gender)) + scale_fill_startrek() + theme_modern()

#Create a plot for SWL by Gender with a different color palette

ggplot(lab3data, aes(x = Gender, y = SWL)) + geom_violin(aes(fill=Gender)) + scale_fill_cosmic() + theme_modern()

```

# Q2: What you can tell from looking at a violin plot? What you cannot tell? What inferences you can make based on the two violin plots? 

The violin plot gives us a visual of the Rosenberg Self-Esteem (RSE plot) and the Satisfaction With Life (SWL plot) frequency distributions, mode(s), and ranges. It can not tell you what individual data points constitute the distribution, the mean, median, standard deviation, standard error, and the number of participants included in the distribution. The inferences we can make about these two plots are that for the RSE plot, male have two peaks and females have one. This suggests that the two most frequently occurring self-esteem level for males tend to be higher than that of females, concentrating at the two points. For the SWL plot, they have very similar distributions, therefore a similar number of women and men have each level of life satisfaction.